---
title: "setup"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

The goal of `gfdata` is to create a clear workflow for pulling and cleaning data for fishery stock assessments based upon the ADMB rockfish assessment framework. 
It utilizes a "project oriented workflow" via RStudio (base R is ok, though you must use `here::here()`). 
You must be able to have a connection to the AFSC & AKFIN (Answers) data servers (e.g., VPN if offsite), and have usernames/passwords setup.

## Installation
The `groundfishr` package can be installed from [github](https://github.com/BenWilliams-NOAA/gfdata) with:
```{r}
# install.packages("devtools")
devtools::install_github("BenWilliams-NOAA/gfdata")
```

```{r setup}
library(gfdata)
```

A consistent directory format is probably the most important aspect of this package. 
The first function creates a number of directories within your R project folder. 
The setup is organized by year within a folder. 
The `setup()` function creates the following folders and files:

```{r}
# setup folders -----

setup(2020)

```
```
project
|__2020
    |__data
        |__models
        |    |__ageage
        |    |     AGEAGE.tpl
        |    |__allometric
        |    |     allometric.tpl
        |    |__length_sd
        |    |     lengthSD.tpl
        |    |__VBL
        |    |      VBL.tpl
        |    |__wVBL
        |    |      wVBL.tpl
        |    |      lvb.ctl
        |
        |__output
        |__raw
        |__SARA
        |__user_input

```

This can then be updated for each year. 

```{r}
# setup folders -----

setup(2021)

```

Which would create a duplication of the folders for the given year

```
project
|__2020
|   |__data
|       |__models
|       |    |__ageage
|       |    |     AGEAGE.tpl
|       |    |__allometric
|       |    |     allometric.tpl
|       |    |__length_sd
|       |    |     lengthSD.tpl
|       |    |__VBL
|       |    |      VBL.tpl
|       |    |__wVBL
|       |    |      wVBL.tpl
|       |    |      lvb.ctl
|       |
|       |__output
|       |__raw
|       |__SARA
|       |__user_input
|__2021
    |__data
        |__models
        |    |__ageage
        |    |     AGEAGE.tpl
        |    |__allometric
        |    |     allometric.tpl
        |    |__length_sd
        |    |     lengthSD.tpl
        |    |__VBL
        |    |      VBL.tpl
        |    |__wVBL
        |    |      wVBL.tpl
        |    |      lvb.ctl
        |
        |__output
        |__raw
        |__SARA
        |__user_input        

```

The SSC accepted model can then be added from the last assessment as the current year's "base model".

```{r}
accepted_model(base_year = 2020, base_model = "m18.2b", year = 2021)
```

This creates a duplication of the folders for the current year

```
project
|__2020
    |__data
    |__base    # last accepted model
    |__m18.2a  # model variant a
    |__m18.2b  # model variant b
|__2021
    |__data
    |__base   # dupication of m18.2b
        |__ all model files

```

Note that if you are fortunate to have a stock that has years with partial assessments the `setup()` function has a switch (`off_yr`). This switch changes the folders/files that are loaded for that year. It provides the `raw` folder for updates to catch data (or design-based survey estimates), an `output` folder for cleaned up catch data, a `SARA` folder for the annual reporting of these data, and the `user_input` folder for storing updates to VAST, etc. that may not be used in the assessment but might make there way into the SAFE. 

```{r}
# setup folders -----

setup(2021, off_yr = TRUE)

```

```
project
|__2020
|   |__data
|       |__models
|       |    |__ageage
|       |    |     AGEAGE.tpl
|       |    |__allometric
|       |    |     allometric.tpl
|       |    |__length_sd
|       |    |     lengthSD.tpl
|       |    |__VBL
|       |    |      VBL.tpl
|       |    |__wVBL
|       |    |      wVBL.tpl
|       |    |      lvb.ctl
|       |
|       |__output
|       |__raw
|       |__SARA
|       |__user_input
|__2021
    |__data
        |__output
        |__raw
        |__SARA
        |__user_input
```
